Medical Concept Extraction

### 2012-12-15 paper/paper.tex
filler crap on stemming do not push

@@

>>>>

The stemmer is divided into a number of linear steps, five or six depending upon the definition of a step, that are used to produce the final stem. The following section will describe each step in turn, combined with a diagram displaying all the key steps of the algorithm. The following definitions regarding the stemmer need to be made before the steps can be explained. A consonant is a letter other than A, E, I, O, U and Y preceded by a consonant. For example the in the word boy the consonants are B and Y, but in try they are T and R. A vowel is any letter that is not a consonant. A list of consonants greater than or equal to length one will be denoted by a C and a similar list of vowels by a V.

Any word can therefore be represented by the single form; [C] (VC)m [V] Where the superscript m denotes m repetitions of VC and the square brackets [] denote the optional presence of their contents [19] The value m is called the measure of a word and can take any value greater than or equal to zero, and is used to decide whether a given suffix should be removed. All such rules are of the form; (condition) S1 -> S2 Which means that the suffix S1 is replaced by S2 if the remaining letters of S1 will satisfy the condition.

The first step of the algorithm is designed to deal with past participles and plurals. This step is the most complex and is separated into three parts in the original definition, 1a, 1b and 1c. The first part deals with plurals, for example sses -> ss and removal of s. The second part removes ed and ing, or performs eed ? ee where appropriate. The second part continues only if ed or ing is removed and transforms the remaining stem to ensure that certain suffices are recognised later. The third part simply transforms a terminal y to an i, this part is inserted as step 2 in the flowchart.

The remaining steps are relatively straightforward and contain rules to deal with different order classes of suffices, initially transforming double suffices to a single suffix and then removing suffices providing the relevant conditions are met.

@@

The stemmer utilises a single table of rules, each of which may specify the removal or replacement of an ending. This technique of replacement is used to avoid the problem spelling exceptions as described earlier, by replacing endings rather than simply removing them the stemmer manages to do without a separate stage in the stemming process, i.e. no recoding or partial matching is required. This helps to maintain the efficiency of the algorithm, whilst still being effective. The rules are indexed by the last letter of the ending to allow efficient searching and are of the following form;

An ending of one or more characters, held in reverse order
An optional intact flag '*'
A digit specifying the removal total (zero or more)
An optional append string of one or more characters
A continuation symbol, '>' or '.'
The following example demonstrate how the rules are stored and used and how replacement can be used to negate the need for recoding. The rule nois4j> causes sion endings to be replaced by j. This acts as a pointer to the j section of the rules, leading to the following transformation; provision -> provij -> provid. The j-transformation is utilised to ensure that the terms provision and provide are correctly conflated to the stem provid.

The algorithm has four main steps detailed below, and presented in the flowchart 
1.Select relevant section; Inspect the final letter of the term and, if present, consider the first rule of the relevant section of the rule table.
2.Check applicability of rule; If final letters of term do not match rule, or intact settings are violated or acceptability conditions are not satisfied go to stage 4.
3.Apply Rule; Remove or reform ending as required and then check termination symbol, and either terminate or return to stage 1.
4.Look for another rule; Move to the next rule in table, if the section letter has changed then terminate, else go to stage 2.
